CHEC Study Sweat Chloride

### 2025-08-05 Edith Zemanick/Sweat Variability/sweat_test_analysis_chec.qmd
CHEC added ethnicity

```{r data cleaning}
# Import CHEC data
chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride.csv",
  na.strings = c("", "NA")
)
# Only include people with either a pre-modulator or post-modulator sodium
chec <- chec %>% filter(!is.na(swsavg), site != "005 Buffalo, NY CC and Peds")
# Calculate ratios
chec$pre_ratio <- chec$swhsavgpm / chec$swhcavgpm
chec$post_ratio <- chec$swsavg / chec$swcavg
# Factor variable formatting
chec$race_txt <- factor(chec$race_txt,
  levels = c(
    "White", "Black or African American", "American Indian or Alaska Native",
    "Native Hawaiian or Other Pacific Islander", "More than One Race",
    "Unknown or Not Reported"
  )
)
chec$modqual <- factor(chec$modqual,
  levels = c(
    "Ivacaftor", "Lumacaftor / Ivacaftor",
    "Tezacaftor / Ivacaftor", "Elexacaftor / Tezacaftor / Ivacaftor"
  ),
  labels = c("IVA", "LUM/IVA", "TEZ/IVA", "ELX/TEZ/IVA")
)
chec$agecat_txt <- factor(chec$agecat_txt,
  levels = c(
    "2-5 years", "6-11 years", "12-17 years", "18-25 years", ">=26 years"
  )
)
# Long dataset for easier plotting and paired analysis
chec_long <- chec %>%
  rename(
    pre_cl = swhcavgpm, pre_na = swhsavgpm, post_cl = swcavg, post_na = swsavg
  ) %>%
  pivot_longer(c(pre_cl, pre_na, post_cl, post_na, pre_ratio, post_ratio),
    names_to = c("pre_post", ".value"),
    names_pattern = "(.*)_(.*)",
    values_to = "count"
  )
chec_long$pre_post <- factor(chec_long$pre_post,
  levels = c("pre", "post"),
  labels = c("Pre-Modulator", "Post-Modulator")
)
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "meansd", "medianrange"),
    cat.stats = c("N", "Nmiss", "countpct"), digits = 1
  )
# Labels
label(chec_long$cl) <- "Sweat Cl"
label(chec_long$na) <- "Sweat Na"
label(chec_long$ratio) <- "Sweat Na:Cl"
```

@@

```{r results='asis'}
t1 <- tableby(
  modqual ~ sex + kwt(age) + agecat_txt +
    race_txt + genotype_group + swcavg + swhcavgpm + swsavg + swhsavgpm +
    pre_ratio + post_ratio,
  data = chec, cat.test = "fe", simulate.p.value = TRUE, control = mycontrols
)
summary(t1, pfootnote = T)
```

- I think that we currently don't have ethnicity data.

>>>>

```{r data cleaning}
# Import CHEC data
chec <- read.csv("./Data_Raw/CHEC sweat sodium and chloride 2024-04-19.csv",
  na.strings = c("", "NA")
)
ethnicity <- read.csv("./Data_Raw/CHEC sweat sodium and chloride 2025-07-30.csv",
  na.strings = c("", "NA")
)
chec <- left_join(chec, ethnicity %>% select(subject, crfethnicity_txt),
  by = join_by(subject)
)
# Only include people with either a pre-modulator or post-modulator sodium
chec <- chec %>% filter(!is.na(swsavg), site != "005 Buffalo, NY CC and Peds")
# Calculate ratios
chec$pre_ratio <- chec$swhsavgpm / chec$swhcavgpm
chec$post_ratio <- chec$swsavg / chec$swcavg
# Factor variable formatting
chec$race_txt <- factor(chec$race_txt,
  levels = c(
    "White", "Black or African American", "American Indian or Alaska Native",
    "Native Hawaiian or Other Pacific Islander", "More than One Race",
    "Unknown or Not Reported"
  )
)
chec$crfethnicity_txt <- factor(chec$crfethnicity_txt,
  levels = c("Not Hispanic or Latino", "Hispanic or Latino")
)
chec$modqual <- factor(chec$modqual,
  levels = c(
    "Ivacaftor", "Lumacaftor / Ivacaftor",
    "Tezacaftor / Ivacaftor", "Elexacaftor / Tezacaftor / Ivacaftor"
  ),
  labels = c("IVA", "LUM/IVA", "TEZ/IVA", "ELX/TEZ/IVA")
)
chec$agecat_txt <- factor(chec$agecat_txt,
  levels = c(
    "2-5 years", "6-11 years", "12-17 years", "18-25 years", ">=26 years"
  )
)
# Long dataset for easier plotting and paired analysis
chec_long <- chec %>%
  rename(
    pre_cl = swhcavgpm, pre_na = swhsavgpm, post_cl = swcavg, post_na = swsavg
  ) %>%
  pivot_longer(c(pre_cl, pre_na, post_cl, post_na, pre_ratio, post_ratio),
    names_to = c("pre_post", ".value"),
    names_pattern = "(.*)_(.*)",
    values_to = "count"
  )
chec_long$pre_post <- factor(chec_long$pre_post,
  levels = c("pre", "post"),
  labels = c("Pre-Modulator", "Post-Modulator")
)
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss", "meansd", "medianrange"),
    cat.stats = c("N", "Nmiss", "countpct"), digits = 1
  )
# Labels
label(chec_long$cl) <- "Sweat Cl"
label(chec_long$na) <- "Sweat Na"
label(chec_long$ratio) <- "Sweat Na:Cl"
```

@@

```{r results='asis'}
t1 <- tableby(
  modqual ~ sex + kwt(age) + agecat_txt +
    race_txt + crfethnicity_txt + genotype_group + swcavg + swhcavgpm + swsavg +
    swhsavgpm + pre_ratio + post_ratio,
  data = chec, cat.test = "fe", simulate.p.value = TRUE, control = mycontrols
)
summary(t1, pfootnote = T)
```
